---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


# wav2textgrid

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/Human-Interaction-Lab/wav2textgrid/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/Human-Interaction-Lab/wav2textgrid/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of `wav2textgrid` is to take a two channel `.wav` file of a conversation and turn it into a transcribed TextGrid ready for further analysis in R or Praat. To use the package, you'll need `R` and `Praat` installed (a python environment is provisioned automatically the first time you run the package).

## Installation

You can install the development version of wav2textgrid from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("Human-Interaction-Lab/wav2textgrid")
```

## Example

### Python Environment Set Up

No manual setup is needed. The package relies on python (for the OpenAI Whisper model); the first time you run it, the package calls

```{r pyrequire, eval=FALSE}
reticulate::py_require("openai-whisper", python_version = "3.11")
```

which makes `reticulate` automatically download and cache a python environment with `openai-whisper` installed.

If you prefer to manage a persistent conda environment yourself, you can still create one with:

```{r example, eval=FALSE}
library(wav2textgrid)
py_setup("wav2textgrid-env")
```

or point `reticulate` at any environment that has `openai-whisper` installed (e.g., `reticulate::use_condaenv()` or `reticulate::use_virtualenv()`) before running `auto_textgrid()`.

### Run `auto_textgrid()`

```{r example1, eval=FALSE}
wav_file = "path/example.wav"
auto_textgrid(wav_file)
```

```{r example2, echo=FALSE, eval=FALSE}
reticulate::py_require("openai-whisper", python_version = "3.11")

library(wav2textgrid)
wav_file = here::here("inst", "extdata", "example.wav")
auto_textgrid(wav_file)
```

```
#> → Default prompt:
#> 'I was like, was like, I'm like, um, ah, huh, and so, so um, uh, and um, mm-hmm, like um, so like, like it's, it's like, i mean, yeah, uh-huh, hmm, right, ok so, uh so, so uh, yeah so, you know, it's uh, uh and, and uh'
#> ✔ Step 1 of 5 [2.6s]
#> ✔ Step 2 of 5 [1.8s]
#> ✔ Step 3 of 5 [5m 11.3s]
#> ✔ Step 4 of 5 [262ms]
#> ✔ Step 5 of 5 [68ms]
#> ℹ Written to path/example_output.TextGrid
```

The output is a textgrid that can be loaded directly into Praat or can be read into R via `readtextgrid` package. For the wav file in the example above (found on GitHub in the `inst` folder), running the code above gives us the following TextGrid (as shown by reading it in using `readtextgrid`).

```{r, echo=FALSE}
text = readtextgrid::read_textgrid(here::here("inst", "extdata", "example_output.TextGrid"))
dplyr::select(text, tier_num, tier_xmin, tier_xmax, xmin, xmax, text)
```

Importantly, we can control a number of features of the transcription to make it more accurate. In this case, we used a noise reduction program and set the threshold of silences to -30 dB. We also use the larger "small" model from Whisper.

```{r, eval=FALSE}
auto_textgrid(wav_file, noise_reduction = TRUE, threshold = -30, model_type = "small")
```

We do recommend the "small" model or larger for more accurate transcription, especially when using the default prompt.
